---
title: "stock analysis"
output:
  pdf_document: default
  html_document: default
editor_options:
  markdown:
    wrap: 120
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(error = TRUE)
knitr::opts_chunk$set(fig.show = 'asis')
knitr::opts_chunk$set(fig.width = 7, fig.height = 7)


# Import libraries
library(gridExtra)
library(knitr)
library(kableExtra)
library(lme4)
library(ICC)
library(knitr)
library(tidyverse)
library(tidyquant)
library(xts)
library(PerformanceAnalytics)
library(corrplot)
library(dplyr)


# Import the CSV file
df <- read.csv("stockdataformatted.csv")

# Identify missing values
sapply(df, function(x) sum(is.na(x)))

# Split the dataframe by the 'Company.Name' column
list_of_companies <- split(df, f = df$Company.Name)

summary(df)


# Initialize a list to hold correlation results for each company
correlations_by_company <- list()




```


```{r}
# Select only the numeric columns
numeric_df <- df %>% select_if(is.numeric)

# Calculate the correlation matrix
cor_matrix <- cor(numeric_df, use = "complete.obs")

# Visualize the correlation matrix
corrplot(cor_matrix, method = "circle")
```
This is likely inaccurate due to a repeating of quarters for every company

```{r}

# Code below gets the correlations for the first 10 companies between last sale price and the rest of the variables


# Get the names of the first 10 companies from the list
first_10_companies <- names(list_of_companies)[1:10]

# Loop through each of the first 10 companies
for(n in first_10_companies) {
  # Select the current company's dataframe
  df <- list_of_companies[[n]]

  # Determine the third-last variable name
  third_last_var <- names(df)[ncol(df)-2]

  # Select only the numeric columns and the third-last variable
  numeric_df <- df %>% select_if(is.numeric) %>% select(third_last_var, everything())

  # Initialize a vector to store the correlations
  cor_vector <- numeric(ncol(numeric_df))
  names(cor_vector) <- names(numeric_df)

  # Calculate the correlation of the third-last variable with all others
  for(column_name in names(numeric_df)) {
    if(sum(complete.cases(numeric_df[[third_last_var]], numeric_df[[column_name]])) > 1) {
      # Compute the correlation if there are enough complete cases
      cor_vector[column_name] <- cor(numeric_df[[third_last_var]], numeric_df[[column_name]], use = "complete.obs")
    } else {
      # Assign NA if not enough complete cases
      cor_vector[column_name] <- NA
    }
  }

  # Check if there are any non-NA values to plot
  if(all(is.na(cor_vector))) {
    cat(paste("Correlation plot for", n, "cannot be generated due to insufficient data.\n"))
  } else {
    # Construct a matrix from the vector for plotting
    cor_matrix <- matrix(cor_vector, nrow = 1, byrow = TRUE)
    colnames(cor_matrix) <- names(cor_vector)
    rownames(cor_matrix) <- third_last_var

    # Plot title with the company name for clarity
    plot_title <- paste("Correlation with", third_last_var, "for", n)

    # Visualize the correlation matrix using corrplot
    print(corrplot(cor_matrix, method = "circle", title = plot_title, is.corr = FALSE))
  }
}
```
As you can see here each company depends on different factors which affect its price

```{r}
# Code gets the mean correlation matrix between last sale price and all other variables


# Initialize a list to store the correlation data for each company
cor_data_list <- list()

# Get the names of the first 500 companies from the list
all_companies <- names(list_of_companies)

# Utility function to check if there's sufficient data to compute the standard deviation
is_var_constant <- function(x) {
  length(unique(na.omit(x))) <= 1
}

# Loop through each of the companies
for(n in all_companies) {
  # Check if the company's data frame exists and is not NULL
  if (!is.null(list_of_companies[[n]]) && ncol(list_of_companies[[n]]) > 2) {
    # Select the current company's dataframe
    df <- list_of_companies[[n]]

    # Determine the third-last variable name
    third_last_var <- names(df)[ncol(df)-2]

    # Select only the numeric columns
    numeric_df <- df %>% select_if(is.numeric)

    # Initialize a vector to store the correlations for the current company
    cor_vector <- rep(NA, ncol(numeric_df))

    # Calculate the correlation of the third-last variable with all other numeric variables
    for(i in seq_along(numeric_df)) {
      # Check if the variable is constant (no variance)
      if (!is_var_constant(numeric_df[[third_last_var]]) && !is_var_constant(numeric_df[[i]])) {
        cor_vector[i] <- cor(numeric_df[[third_last_var]], numeric_df[[i]], use = "complete.obs")
      }
      # Else, the result is already NA
    }

    # Name the elements of the vector with the column names
    names(cor_vector) <- names(numeric_df)

    # Store the results in the cor_data_list
    cor_data_list[[n]] <- cor_vector
  } else {
    # Print a message if the company's data frame does not exist or is NULL
    cat(paste("The data frame for company", n, "does not exist or is NULL.\n"))
  }
}



  # Name the elements of the vector with the column names
  names(cor_vector) <- names(numeric_df)

  # Store the results in the cor_data_list
  cor_data_list[[n]] <- cor_vector

# Combine all the correlation vectors into one data frame
cor_data_frame <- do.call(rbind, cor_data_list)

# Compute the mean of the correlation coefficients for the third-last variable across all companies
mean_correlation <- colMeans(cor_data_frame, na.rm = TRUE)

# Output the mean correlation as a single row data frame
mean_correlation_df <- as.data.frame(t(mean_correlation))

corrplot(as.matrix(mean_correlation_df), is.corr=FALSE)
```
On average most variables are correlated with price, but many of these variables are also correlated with each other

```{r}
# Calculating the mean full correlation matrix


# Initialize a list to store the full correlation matrix for each company
full_cor_matrices <- list()
common_columns <- NULL


# Loop through each company
for(n in all_companies) {
  df <- list_of_companies[[n]]
  if (!is.null(df) && ncol(df) > 2) {
    numeric_df <- df %>%
      select_if(is.numeric) %>%
      na.omit()

    # Check if there's enough variation in each column to calculate correlation
    numeric_df <- numeric_df %>%
      select_if(~sd(.) != 0)

    if(nrow(numeric_df) >= 2) {
      cor_matrix <- cor(numeric_df)
      full_cor_matrices[[n]] <- cor_matrix
      
      # Identify common columns if not already identified
      if (is.null(common_columns)) {
        common_columns <- names(numeric_df)
      } else {
        common_columns <- intersect(common_columns, names(numeric_df))
      }
    }
  }
}

# Function to compute the mean correlation matrix from the list of matrices
compute_mean_cor_matrix <- function(cor_matrices, common_columns) {
  # Subset matrices to only common columns
  cor_matrices <- lapply(cor_matrices, function(m) m[common_columns, common_columns])
  
  # Calculate mean correlation matrix using Reduce to sum the matrices
  sum_matrix <- Reduce("+", cor_matrices)
  counts <- Reduce("+", lapply(cor_matrices, function(m) !is.na(m)))
  
  # Calculate the mean matrix by dividing sum by counts
  mean_matrix <- sum_matrix / counts
  
  # Replace NaN values with NA
  mean_matrix[is.nan(mean_matrix)] <- NA
  return(mean_matrix)
}

# Ensure we have correlation matrices and common columns to work with
if(length(full_cor_matrices) > 0 && !is.null(common_columns)) {
  mean_cor_matrix <- compute_mean_cor_matrix(full_cor_matrices, common_columns)
  
  # Visualize the mean correlation matrix
  corrplot(mean_cor_matrix, method = "circle", title = "Mean Correlation Matrix Across All Companies")
} else {
  cat("No correlation matrices available to compute the mean matrix.\n")
}

```
This is way more accurate than the full correlation matrix from earlier

```{r}

# Does the same thing as the mean correlation matrix between last sale price and all other variables but groups by industry

# Initialize a list to store the correlation data for each company
cor_data_list <- list()

# Get the names of the first 500 companies from the list
all_companies <- names(list_of_companies)

# Utility function to check if there's sufficient data to compute the standard deviation
is_var_constant <- function(x) {
  length(unique(na.omit(x))) <= 1
}

# Loop through each of the companies
for(n in all_companies) {
  # Check if the company's data frame exists and is not NULL
  if (!is.null(list_of_companies[[n]]) && ncol(list_of_companies[[n]]) > 2) {
    # Select the current company's dataframe
    df <- list_of_companies[[n]]

    # Determine the third-last variable name
    third_last_var <- names(df)[ncol(df)-2]

    # Select only the numeric columns
    numeric_df <- df %>% select_if(is.numeric)

    # Initialize a vector to store the correlations for the current company
    cor_vector <- rep(NA, ncol(numeric_df))

    # Calculate the correlation of the third-last variable with all other numeric variables
    for(i in seq_along(numeric_df)) {
      # Check if the variable is constant (no variance)
      if (!is_var_constant(numeric_df[[third_last_var]]) && !is_var_constant(numeric_df[[i]])) {
        cor_vector[i] <- cor(numeric_df[[third_last_var]], numeric_df[[i]], use = "complete.obs")
      }
      
      # Else, the result is already NA
    }

    # Name the elements of the vector with the column names
    names(cor_vector) <- names(numeric_df)
    
    # Inside the loop, after calculating cor_vector for each company:
    cor_vector['Primary.Industry'] <- df$Primary.Industry[1]


    # Store the results in the cor_data_list
    cor_data_list[[n]] <- cor_vector
  } else {
    # Print a message if the company's data frame does not exist or is NULL
    cat(paste("The data frame for company", n, "does not exist or is NULL.\n"))
  }
}


  # Name the elements of the vector with the column names
  names(cor_vector) <- names(numeric_df)

  # Store the results in the cor_data_list
  cor_data_list[[n]] <- cor_vector
  
# Convert cor_data_list to a data frame
cor_data_frame_with_industry <- do.call(rbind, cor_data_list)

# Convert to a proper data frame and handle factor conversion for the 'Primary.Industry' column
cor_data_frame_with_industry <- data.frame(cor_data_frame_with_industry, stringsAsFactors = FALSE)
cor_data_frame_with_industry$Primary.Industry <- as.factor(cor_data_frame_with_industry$Primary.Industry)

# Calculate the mean correlation by industry
mean_correlation_by_industry <- cor_data_frame_with_industry %>%
  group_by(Primary.Industry)


library(dplyr)
library(tibble)

industry_filtered <- na.omit(mean_correlation_by_industry) 


num_cols <- ncol(industry_filtered) - 1  # Get the number of columns minus 1
industry_filtered[, 1:num_cols] <- lapply(industry_filtered[, 1:num_cols], as.numeric) 

industry_correlations <- industry_filtered %>%
  group_by(Primary.Industry) %>%
  summarize(across(where(is.numeric), mean))

head(industry_correlations, 10)

```
Depending on the industry, each variable can be strongly or negatively correlated with last sale price


```{r}
# Model Building

# Level 1 - Individual time period observations

# Every level 1 variable has a strong correlation with at least one industry, so it's best to leave all variables in the model other than Year, but highly intercorrelated variable pairs such as EBITDA and EBITDA margin you can just select one of them



# Level 2 - Company (Grouping Factor)

# Primary.Industry and Company Name are the only level 2 covariates


```



